---
title: "AE04 - Getting to Know Your Data Types"
author: "Anthony Palomba"
date: "2026-09-24"
format: html
execute:
  warning: false
  message: false
---

## Setup

```{r}
#| label: packages
library(tidyverse)
library(googlesheets4)
```

## Step 1: Import

```{r}
#| label: import

raw <- read_sheet("https://docs.google.com/spreadsheets/d/1e8XSidxeT3SIE10g1jiEuALRa7REpmYU64_3V3TNKKs/edit?usp=sharing") %>%
  rename(
    timestamp       = Timestamp,
    zipcode         = `What is the zipcode of your hometown?`,
    stats_before    = `Have you taken any statistics courses before?`,
    prog_experience = `How much experience do you have with programming?`,
    languages       = `If you've programmed before, which languages have you used, and how comfortable do you feel with each? If you haven't programmed before, please leave this question blank.`,
    data_interest   = `What types of data interest you?`,
    learn_best      = `How do you learn best?`,
    commute_min     = `How long does it take you to commute from home to this classroom, in minutes? Please enter a number only.`
  )
```

## Step 2: Data type audit

#### *"How long does it take you to commute from home to this classroom, in minutes? Please enter a number only."*

```{r}
#| label: type-check-commute
class(raw$commute_min)

```

Is that what you'd expect, given the question?

No. I would expect an integer/number.

```{r}
#| label: commute-mean-broken
mean(raw$commute_min)
```

What's going on? `unlist()` this variable and look at the `unique()` values.

```{r}
#| label: commute-unique
unlist(raw$commute_min)
```

#### *"How much experience do you have with programming?"*

```{r}
#| label: type-check-prog
class(raw$prog_experience)
table(raw$prog_experience)
```

Is this variable ordinal? Does everything you'd expect to see actually show up?

#### Skim the rest

```{r}
#| label: skim-rest
glimpse(raw)
```

For each of the other variables — is R's type what you'd expect, given the question that was asked? We won't work with them further in this AE, but you should still be able to answer this for each one.

## Step 3: Apply fixes

#### Commute time

While we could try to coerce list → character → numeric, it's easier here to specify data types on read.

```{r}
#| label: readsurvey-specify-coltypes


raw <- read_sheet("https://docs.google.com/spreadsheets/d/1e8XSidxeT3SIE10g1jiEuALRa7REpmYU64_3V3TNKKs/edit?usp=sharing", col_types = c("Dccccccn")) %>%
  rename(
    timestamp       = Timestamp,
    zipcode         = `What is the zipcode of your hometown?`,
    stats_before    = `Have you taken any statistics courses before?`,
    prog_experience = `How much experience do you have with programming?`,
    languages       = `If you've programmed before, which languages have you used, and how comfortable do you feel with each? If you haven't programmed before, please leave this question blank.`,
    data_interest   = `What types of data interest you?`,
    learn_best      = `How do you learn best?`,
    commute_min     = `How long does it take you to commute from home to this classroom, in minutes? Please enter a number only.`
  )
```

Always keep an eye out for this, and ask yourself whether it's expected, given what you found in Step 2.

```{r}
#| label: commute-values
class(raw$commute_min)
```

#### Recode & reorder `prog_experience`

1.  First, confirm what's actually in the data.

```{r}
#| label: prog-distinct
table(raw$prog_experience)
glimpse(raw$prog_experience)
typeof(raw$prog_experience)
```

2.  Then compare against the published [survey](https://forms.gle/ktGiD3R4dGNabASx7). Is everything you'd expect to see actually there?

> No. I'm missing the category "A little..."

3.  Convert the variable to a factor with the current recorded levels.

```{r}
#| label: prog-recode
raw$prog_experience <- factor(raw$prog_experience, levels = c("None", "Some — I’ve worked on a few projects or used it occasionally", "A lot — I use programming regularly and feel confident writing code"))

```

4.  Now that it's reading as a factor, use `fct_expand()` to add any level that's missing.

```{r}
#| label: prog-expand
fct_expand(raw$prog_experience, "A little — I’ve written a few lines or done small exercises")
```

5.  Last, respect the ordinal nature of this variable by putting the levels in increasing order.

```{r}
#| label: prog-relevel
raw$prog_experience <- fct_relevel(raw$prog_experience, "A little — I’ve written a few lines or done small exercises", after = 1)
table(raw$prog_experience)
```
